### 2017-10-26 src/4_OnlineLearning.tex
Adding appendix chapter. page 49

\chapter{Machine Learning Models}
\label{chapter:MLM}
Machine learning is a scientific discipline focused on the development of
algorithms based on learning from examples. This idea has become central to the
design of search engines, robots systems and forecasts applications that process
large data sets. Usually machines designed to forecast financial time series
requires long training period and therefore they are not suitable to be updated
with stream data. Online machine learning techniques tackle this problem by
updating the model with new data or by computing a new model using less data so
they can give a response in a short period of time.

@@

Regularization can be understood using two contexts: learning theory
(probabilistic) and inverse problems (deterministic). In the context of
learning, regularization refers to techniques allowing to avoid over-fitting and
the desire property of the selected estimator is to perform well on new data (to
generalize), e.g. regularized least squares. In the context of inverse problems,
regularization objective is to stabilise, with respect to noise, a possibly
ill-conditioned matrix inversion problem e.g spectral cut-off and Tikhonov
regularization or ridge regression (RR) \cite{ tikhonov1977}.

@@

\begin{equation}
\label{eq:RRdefinitions}
\mathbf{X} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{x}_{1}    & \mathbf{x}_{2}    & \ldots & \mathbf{x}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\quad , \quad
\mathbf{A} = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    --- & \mathbf{a}^{\top}_{1} & ---\\
    --- & \mathbf{a}^{\top}_{2} & ---\\
    & \vdots & \\
    --- & \mathbf{a}^{\top}_{m} & ---
  \end{tabular}
\right]
\quad \text{and} \quad
\mathbf{Y} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{y}_{1}    & \mathbf{y}_{2}    & \ldots & \mathbf{y}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\end{equation}

@@

\noindent where $f(\mathbf{x}_t)={\mathbf{A}}^\top {\mathbf{x}}_t$. To solve equation \ref{eq:lsm} is equivalent to find a solution of:
\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\mathbf{X}}} = \underset{m \times l}{\mathbf{Y}}
\end{equation}

@@

\textbf{Proof}\quad

In the case matrix $\mathbf{A}$ is non singular, its solution is
straight forward:

\begin{equation}
\label{OLSsolution}
    \mathbf{\mathbf{X}}=\mathbf{A}^{-1}\mathbf{Y}
\end{equation}

Since the problem shown in equation~(\ref{eq:regproblem}) has not
solution, the minimum norm given by equation~(\ref{OLSsolution}) is
obtained by solving the equivalent problem:

\begin{equation*}
\label{eq:proyectorsol}
\mathbf{A \hat{\mathbf{X}} = PY} 
\end{equation*}

\noindent where $\mathbf{P=U_1 U_1^\top}$ is the projection onto the
Col($\mathbf{A}$).

Since $\mathbf{V} = [\underset{(n \times k)}{\mathbf{V_1}} |
\underset{(n \times k)}{\mathbf{V_2}}]$ and $\mathbf{V_1^\top V_2 =
0}$ we can express $\mathbf{\hat{\mathbf{X}}} = \mathbf{V_1 \mathbf{X}_1 + V_2 \mathbf{X}_2}$
with $\mathbf{\mathbf{X}_2=0}$ because $\mathbf{\hat{\mathbf{X}}}$ lives in the
$\text{Row}(\mathbf{A})$ given by $\mathbf{V_1}$, so we have:

\begin{eqnarray*}
\mathbf{A \hat{\mathbf{X}}} &=& \mathbf{PY} \\
\mathbf{U_1 \Sigma_1 V_1^\top \hat{\mathbf{X}}} &=& \mathbf{U_1 U_1^\top Y} \\
\mathbf{ V_1^\top \hat{\mathbf{X}}} &=&  \mathbf{\Sigma_1^{-1} U_1^\top Y} \\ 
\mathbf{ V_1^\top V_1 \mathbf{X}_1} &=& \mathbf{\Sigma_1^{-1}
U_1^\top Y} \\
\mathbf{\mathbf{X}_1}&=& \mathbf{\Sigma_1^{-1} U_1^\top Y}
\end{eqnarray*}

\noindent from this result we can obtain $\mathbf{\hat{\mathbf{X}}}$ and
therefore the pseudo-inverse expression:

\begin{eqnarray*}
\mathbf{\hat{\mathbf{X}}} &=& \mathbf{V_1 \mathbf{X}_1} \\
                &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top Y} \\
\mathbf{A^+} &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top} \, .
\end{eqnarray*}

$\blacksquare$

@@

\textbf{Proof}\quad

\begin{eqnarray*}
J(\mathbf{\mathbf{X}}) &=&  \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2  + \lambda
 \| \mathbf{\mathbf{X}}\| ^2 \\
 &=&  \sum_{t=1}^N (\mathbf{\mathbf{X}}^\top {\bf a}_t-y_t)^2 + \lambda \sum_{i=1}^p \mathbf{X}_i^2 \\
 &=& (\mathbf{X}^\top a_1-y_1)^2 + \dots + (\mathbf{X}^\top a_N-y_N)^2 + \lambda (\mathbf{X}_1^2+\dots+\mathbf{X}_p^2)
\end{eqnarray*}
\noindent taking derivatives
 \begin{eqnarray*}
 \frac{\partial J(\mathbf{\mathbf{X}})}{\partial \mathbf{X}_1}&=& 
 2(\mathbf{X}^\top \mathbf{a}_1-y_1)\mathbf{a}_{11} + \dots + 2(\mathbf{X}^\top \mathbf{a}_N-y_N)\mathbf{a}_{N1} + 2\lambda \mathbf{X}_1 \\
 &=& 2\mathbf{a}_1^\top(\mathbf{A}\mathbf{X}-\mathbf{Y}) + 2\lambda\mathbf{X}_1\\
& \vdots &\\
  \frac{\partial J(\mathbf{\mathbf{X}})}{\partial \mathbf{X}_p}&=& 
 2\mathbf{a}_p^\top(\mathbf{A}\mathbf{X}-\mathbf{Y}) + 2\lambda\mathbf{X}_p 
\end{eqnarray*}

Then we have that:
\begin{eqnarray*}
\frac{\partial J(\mathbf{\mathbf{X}})}{\partial \mathbf{X}}&=& 
 2\mathbf{A}^\top(\mathbf{A}\mathbf{X}-\mathbf{Y}) + 2\lambda\mathbf{X}
\end{eqnarray*}
Since $\frac{\partial J(\mathbf{\mathbf{X}})}{\partial \mathbf{X}}=0$ we have:
\begin{eqnarray*}
2\mathbf{A}^\top(\mathbf{A}\mathbf{X}-\mathbf{Y}) + 2\lambda\mathbf{X}&=&0 \\
\mathbf{A}^\top\mathbf{A}\mathbf{X} - \mathbf{A}^\top\mathbf{Y} + \lambda\mathbf{X} &=& 0\\
(\mathbf{A}^\top\mathbf{A}+\lambda\mathbb{I})\mathbf{X} &=&  \mathbf{A}^\top\mathbf{Y} \\
\mathbf{X} &=& (\mathbf{A}^\top\mathbf{A}+\lambda\mathbb{I})^{-1}  \mathbf{A}^\top\mathbf{Y}
\end{eqnarray*}

$\blacksquare$

The method is called ridge regression because the term $\lambda \mathbb{I}$ adds positive entries along the diagonal (ridge) to avoid the
singularity of the covariance matrix $\mathbf{A}^\top \mathbf{A}$. This addition ensures that all of the covariance matrix eigenvalues will be strictly greater than 0, i.e the solution becomes unique.
%=======================HASTA ACA REVISADO PROFE =========================

@@

The solution of OLS is well known $\hat{\mathbf{X}}=(\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y}$, and its bias is:

\begin{eqnarray*}
Bias(\hat{f}(\hat{\mathbf{X}})) &=& E[\hat{\mathbf{X}}] - \mathbf{X} \\
&=& E[ (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y}] - \mathbf{X} \\
&=& E[ (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top (\mathbf{AX})] - \mathbf{X}  \\
&=& \mathbf{X}  - \mathbf{X}  \\
&=&  0
\end{eqnarray*}

@@

\begin{eqnarray*}
\mathbf{X}(\lambda) &=&( \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1}\mathbf{A}^\top \mathbf{Y} \\
&=& (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y} \\
&=&  (\mathbb{I} + \lambda (\mathbf{A}^\top \mathbf{A})^{-1})^{-1}  \hat{\mathbf{X}} \\
&=& \mathbf{W} \hat{\mathbf{X}} 
\end{eqnarray*}

\noindent where $\mathbf{W}  = (\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1}  $ it is defined for simplicity. Ridge
regression bias is then obtained as:

\begin{eqnarray*}
Bias(\mathbf{X}(\lambda)) &=& E[\mathbf{X}(\lambda)] - \mathbf{X} \\
&=& E[\mathbf{W}\hat{\mathbf{X}}] - \mathbf{X} \\
&=&  \mathbf{W} \mathbf{X} - \mathbf{X} \neq 0 
\end{eqnarray*}

@@

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& Var(\mathbf{W}\hat{\mathbf{X}}) \\
&=& E[(\mathbf{W}\hat{\mathbf{X}}-E[\mathbf{W}\hat{\mathbf{X}}])(\mathbf{W}\hat{\mathbf{X}}-E[\mathbf{W}\hat{\mathbf{X}}])^\top] \\
&=& \mathbf{W}E[(\hat{\mathbf{X}}-E[\hat{\mathbf{X}}])(\hat{\mathbf{X}}-E[\hat{\mathbf{X}}])^\top] \mathbf{W}^\top \\
&=& \mathbf{W}Var(\hat{\mathbf{X}})\mathbf{W}^\top \\
&=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top
\end{eqnarray*}

@@

Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}. Ridge regression shows an
increasing squared bias and a decreasing variance.

\textbf{Proof}\quad

Since $Bias(\mathbf{X}(\lambda)) \neq 0$ this imply that

\begin{equation*}
Bias(\mathbf{X}(\lambda)) ^2 > 0 
\end{equation*}

\noindent we know that $\mathbf{A}^\top
\mathbf{A}$ has an eigenvalue decomposition $\mathbf{A}^\top
\mathbf{A} = \mathbf{V} \Sigma \mathbf{V}^{-1}$.

\begin{eqnarray*}
Var(\mathbf{X}(\lambda)) &=& \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top\\
Var(\mathbf{X}(\lambda)) &=& \sigma^2 (\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1} (\mathbf{A}^\top \mathbf{A} )^{-1}((\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1} )^\top \\
&=& \sigma^2 (\mathbb{I} + \lambda (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1} (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1}((\mathbb{I} + \lambda (\mathbf{V} \Sigma \mathbf{V}^{-1})^{-1})^{-1} )^\top \\
&=& \sigma^2 \mathbf{V} (\mathbb{I} + \lambda \Sigma^{-1})^{-1} \Sigma^{-1} (\mathbb{I} + \lambda \Sigma^{-1})^{-1}  \mathbf{V}^{-1}\\
&=& \sigma^2 \mathbf{V} \mathbf{D}^{-1} \mathbf{V}^{-1}
\end{eqnarray*}

\noindent where $\mathbf{D}^{-1}  = (\mathbb{I} + \lambda \Sigma^{-1})^{-1} \Sigma^{-1} (\mathbb{I} + \lambda \Sigma^{-1})^{-1}$ is a diagonal matrix with coefficients:

\begin{equation*}
d_i = \frac{\sigma_i^{-1}}{(1+\lambda\sigma_i^{-1})^2}
\end{equation*}

$\blacksquare$

>>>>

\chapter{Machine Learning Models}
\label{chapter:MLM}
Machine learning is a scientific discipline focused on the development of
algorithms based on learning from examples. This idea has become central to the
design of search engines, robots systems and forecasts applications that process
large data sets. Usually machines designed to forecast financial time series
requires long training period and therefore they are not suitable to be updated
with stream data. Online machine learning techniques tackle this problem by
updating the model with new data or by computing a new model considering less data so
they can give a response in a short period of time.

@@

Regularization can be understood using two contexts: learning theory
(probabilistic) and inverse problems (deterministic). In the context of
learning, regularization refers to techniques allowing to avoid over-fitting and
the desire property of the selected estimator is to perform well on new data (to
generalise), e.g. regularized least squares. In the context of inverse problems,
regularization objective is to stabilise, with respect to noise, a possibly
ill-conditioned matrix inversion problem e.g spectral cut-off and Tikhonov
regularization or ridge regression (RR) \cite{ tikhonov1977}.

@@

@@

\noindent where 
\begin{equation}
\label{eq:RRdefinitions}
\mathbf{X} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{x}_{1}    & \mathbf{x}_{2}    & \ldots & \mathbf{x}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\quad , \quad
\mathbf{A} = 
\left[
  \begin{tabular}{c>{$}c<{$}c}
    --- & \mathbf{a}^{\top}_{1} & ---\\
    --- & \mathbf{a}^{\top}_{2} & ---\\
    & \vdots & \\
    --- & \mathbf{a}^{\top}_{m} & ---
  \end{tabular}
\right]
\quad \text{and} \quad
\mathbf{Y} = 
\left[
  \begin{array}{cccc}
    \vertbar & \vertbar &        & \vertbar \\
    \mathbf{y}_{1}    & \mathbf{y}_{2}    & \ldots & \mathbf{y}_{n}    \\
    \vertbar & \vertbar &        & \vertbar 
  \end{array}
\right]
\end{equation}

\noindent and $f(\mathbf{x}_t)={\mathbf{A}}^\top {\mathbf{x}}_t$.

To solve equation \ref{eq:lsm} is equivalent to find a solution of:
\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\mathbf{X}}} = \underset{m \times l}{\mathbf{Y}}
\end{equation}

@@

Proof can be found in the appendix section \ref{app:pseudoproof}.

@@

Proof can be found in section \ref{app:rroptsection}.

The method is called ridge regression because the term $\lambda \mathbb{I}$ adds positive entries along the diagonal (ridge) to avoid the
singularity of the covariance matrix $\mathbf{A}^\top \mathbf{A}$. This addition ensures that all of the covariance matrix eigenvalues will be strictly greater than 0, i.e the solution becomes unique.

@@

The solution of OLS is well known $\hat{\mathbf{X}}=(\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \mathbf{Y}$, and its bias ($Bias(\hat{f}(\hat{\mathbf{X}})$)  is 0 (Proof in section \ref{app:biasandvariance}).

@@

\begin{equation}
 Bias(\mathbf{X}(\lambda))  = \mathbf{W} \mathbf{X} - \mathbf{X} \neq 0 
 \end{equation}

 \noindent where $\mathbf{W}  = (\mathbb{I} + \lambda (\mathbf{A}^\top
\mathbf{A})^{-1})^{-1}  $

@@

\begin{equation*}
Var(\mathbf{X}(\lambda)) = \sigma^2 \mathbf{W}(\mathbf{A}^\top \mathbf{A} )^{-1}\mathbf{W}^\top
\end{equation*}

Proof in section \ref{app:biasandvariance}.

@@

Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}. Ridge regression shows an
increasing squared bias and a decreasing variance. See proof in section \ref{app:rrbiasvar}.
